---
title: 'Week 1: Recap'
author: "Hong Xiang Yue"
date: "07/03/2019"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
#Lecture Refresher

## Predictive vs Prescriptive Analytics

###Predictive analytics
- Uses variables to predict outcome
- Only cares about $\textit{correlation}$ 
- Isn't interested in changing outcome, but predicting what the outcome will be, so strategies can be used to exploit it
- e.g. I want to predict tomorrow's stock price so that I know whether to buy or sell shares. I'm not interested in what actually caused the stock price change.

###Prescriptive analytics
- Looks at underlying causal relationship between $x$ and $y$
- Not just correlation, but $\textit{causality}$
- Making a $\textit{prescription}$ on how to change outcome, very useful in economic policy 
- e.g. Does forcing people to stay in school for longer increase their wage?


##Types of data
- At a disadvantage when compared to most sciences, don't usually have access to experimental data, only observational
- Very difficult to draw conclusions about causality, need to be very careful
- Cross-sectional: variablesmultiple indivdiuals/firms/countires at a single point in time
- Time-series: observe variables for one entity across time
- Panel: cross-sectional data, but obvserved across time
- Cross-sectional vs time-series: unordered vs ordered, independent vs dependent

##Modelling
- Where theories from economics and finance can be quantified using data
- Regression models will be used heavily in this unit

#Tutorial refresher

##EViews
- Viewing variables
- Histograms and descriptive statistics 
- Correlations and scatter plots
- Linear regression and prediction
- How to save output into Word

##Matrix algebra
- How matrices can be used to simplify our regression equations
- Rows vs columns of a matrix
- Transpose operator
$$
\begin{bmatrix}
1 & 2\\
3 & 4\\
5 & 6\\
\end{bmatrix}'
=
\begin{bmatrix}
1 & 3 & 5\\
2 & 4 & 6
\end{bmatrix}
$$
- How to multiply matrices

#Week 2 Tutorial: Part A

##Vlookup
- Formula to search a table for a piece of information 
- First specify the lookup value, this will be usually some sort of code which identifies a particular individual, country, product, stock etc.
- Then select the table in which you wish to conduct the lookup, the item code should always be the first column of the table
- Next, select the variable you are interested in, this is given by the column number of the table
- Last argument, set as FALSE for our purposes

#Week 2 Tutorial: Part B

##Question 1

###1. 
- Open up your Part A excel file in EViews
- Obtain summary statistics and histogram of GDPPC, in groups of 4 discuss what you notice
- If a different set of countries were sampled would the summary statistics be the same? 
- https://flux.qa/LDPPHD


##Question 1
###2. 
- Create scatterplot between co2pc and gdppc
- What sort of relationship is there?
- Use a log transformation on gdppc and draw a new scatterplot, what do you see?

##Question 2
- Visit www.ausmacrodata.org and download the wage price index csv
- Keep first two columns and delete everything else, open up in Eviews

###1. 
- Plot as line graph
- What do we learn from the plot?

###2. 
- Given this data set, what do you think is the best way to predict next quarter's wage price index?
- https://flux.qa/LDPPHD


##Question 2

###3.
- Very difficult to see other features of the time series because the trend is very strong
- Can look at growth rates instead
- Today's growth rate is 
$$g_t = 100 \times \frac{wpi_t-wpi_{t-1}}{wpi_{t-1}}$$
- e.g. Share price yesterday is 50, today it is 52, the growth rate (as a percentage) is 
$$100\times\frac{52-50}{50}=4\%$$
- Known as the simple return

##Question 2
###3. continued

- Can also calculate log return 
$$g_t =100 \times\Delta \log(wpi_t) =  100\times(\log(wpi_t)-\log(wpi_{t-1}))$$
- Using the same example, the log return is 
$$100\times(\log(52)-\log(50))=3.922071\%$$

- In EViews type wpi_log_returns = dlog(wage) to create a series of log returns on the wage price index
- Open up the new series, why is the first value NA?

##Question 2
###4. 
- Plot the series as a line graph and hover your mouse over the peaks
- What do you notice?

###5.
- Now instead of selecting "line and symbol" select the specific option as "seasonal plot"
- There are two further options in the seasonal type box, "Paneled lines & means" and "Multiple overlayed lines", try both
- Comparing the two plots, what do you see?

##Question 2
###6. Discussion
- Is this bad?
- What other information do we need?

#Part C: How to use MoVE

##The basics
- Install Citrix receiver
- Login to MoVE
- Open and edit files in EViews
- Import files locally

#Part D: More matrix algebra

##Inverse of a matrix
- Note, we can only multiply matrices together, not divide them
- Instead, what we do is multiply by its inverse
- The inverse is only defined for square matrices

$$\textbf{A}^{-1}\textbf{A}=\textbf{I}$$
An example, solving for the matrix $\textbf{A}$

$$\textbf{BA}=\textbf{C}$$
$$\textbf{B}^{-1}\textbf{BA}=\textbf{B}^{-1}\textbf{C}$$
$$\textbf{I}\textbf{A}=\textbf{B}^{-1}\textbf{C}$$
$$\textbf{A}=\textbf{B}^{-1}\textbf{C}$$

##Drawing vectors
```{r}
par(pty="s")

x = c(0,2,-1)
y = c(0,1,2)
      
vdata = data.frame(x,y)

x = as.list(x)
y = as.list(y)

for (i in 1:3){
  x[i] <- toString(x[i])
  y[i] <- toString(y[i])
}


rownames(vdata) = c(paste("(",x[1], ",", y[1], ")"),
                    paste("(",x[2], ",", y[2], ")"),
                    paste("(",x[3], ",", y[3], ")"))
                    


plot(y~x, 
     ylim = c(-1,3),
     xlim = c(-2,3),
     main = "", xlab = "x", ylab = "y",
     data = vdata)

with(vdata, text(y~x, labels = rownames(vdata), pos = 4))


arrows(0,0,vdata$x[2],vdata$y[2], lwd=2)
arrows(0,0,vdata$x[3],vdata$y[3], lwd=2)
```

##Orthogonality 
- Two vectors are orthogonal, if they are perpendicular
- Difficult to visualise in more than three dimensions, but can be verified mathematically 
- Given two column vectors $\textbf{u}$ and $\textbf{v}$, if $\textbf{u}'\textbf{v} = 0$, then the vectors are orthogonal

###Example
$$
\textbf{u} =
\begin{bmatrix}
-1\\
2\\
\end{bmatrix}
,
\textbf{v} =
\begin{bmatrix}
2\\
1\\
\end{bmatrix}
$$
$$\textbf{u}'\textbf{v} = -1\times2 + 2\times 1 = 0$$

##Linear combinations of vectors
Vectors of the same dimension can also be combined to create another vector of the same dimension, by adding and subtracting multiples of given vectors e.g.

$$
\begin{bmatrix}
-1\\
4\\
3
\end{bmatrix}
=
2\times
\begin{bmatrix}
1\\
2\\
3
\end{bmatrix}
-3
\times
\begin{bmatrix}
1\\
0\\
1
\end{bmatrix}
$$
More generally, 

$$\textbf{v} = \sum_{j=1}^k a_j \textbf{u}_j$$
Where the $a_j$ are scalar constants and the $\textbf{u}_j$ are vectors of the same dimension.

##Exercises

###Question 1.
$\newline$
Given 

$$
\textbf{A}
=
\begin{bmatrix}
-4 & 0\\
0 & 9\\
3 & 0 
\end{bmatrix}, 
\textbf{v}
=
\begin{bmatrix}
3 \\
0 \\
4
\end{bmatrix}
$$
Show that each of the columns of $\textbf{A}$ are orthogonal to $\textbf{v}$

##Exercises

###Question 2.
$\newline$
Substitute
$$\hat{\textbf{u}}=\textbf{y}-\textbf{X}\hat{\boldsymbol{\beta}}$$
into 

$$\textbf{X}'\hat{\textbf{u}}=\textbf{0}$$

and solve for $\hat{\boldsymbol{\beta}}$.

##Exercises

###Question 3.
Given 

$$
\textbf{v}=
\begin{bmatrix}
2\\
1\\
-10
\end{bmatrix},
\textbf{u}_1=
\begin{bmatrix}
2\\
0\\
-3
\end{bmatrix},
\textbf{u}_2=
\begin{bmatrix}
2\\
-1\\
4
\end{bmatrix}
$$
Show that $\textbf{v}$ is a linear combination of $\textbf{u}_1$ and $\textbf{u}_2$ i.e. solve for $a_1$ and $a_2$ so that 

$$\textbf{v} =  a_1\textbf{u}_1+a_2\textbf{u}_2$$

is true.
